\class : Week \psetnum

### 2022-11-18 541_ism/week7/solutions.tex
finish week 7 problems for ISM

>>>>

\question{\textbf{2. Broadening}}

\question{2a. Optically Thin Solution}
\answer{
    We have a system of two equations with two unknowns, let's do some rearranging. I'm also going to refer to the FWHM as $w$ for brevity.
    \begin{align}
        \frac{w_1^2}{8 \ln 2} &= \sigma_{v, {\rm turb}}^2 + \frac{k_B T}{M_1},\qquad \frac{w_2^2}{8 \ln 2} = \sigma_{v, {\rm turb}}^2 + \frac{k_B T}{M_2} \\
        \frac{w_1^2}{8 \ln 2} - \frac{k_B T}{M_1} &= \frac{w_2^2}{8 \ln 2} - \frac{k_B T}{M_2} \\
        \frac{w_1^2}{8 \ln 2} - \frac{w_2^2}{8 \ln 2} &= T \qty(\frac{k_B}{M_1} - \frac{k_B}{M_2}) \\
        T &= \frac{w_1^2 - w_2^2}{8 \ln 2} \qty(\frac{k_B}{M_1} - \frac{k_B}{M_2})^{-1}
    \end{align}
    We can plug in $w_1 = 10 \unit{km}{s^{-1}}$, $w_2 = 5 \unit{km}{s^{-1}}$, $M_1 = 1 \unit{amu}$ and $M_2 = 23 \unit{amu}$ to find that the value for an optically thin solution are
    \begin{equation}
        \boxed{ T = 1713 \unit{K},\qquad \sigma_{v, {\rm turb}} = 1.97 \unit{km}{s^{-1}} }
    \end{equation}
}

\question{2b. What if it's not optically thin?}
\answer{
    The FHWM is a larger when it is saturated and so this represents \textbf{an upper limit on the FWHM}.
    \begin{equation}
        \sigma_{v, {\rm turb}}^2 + \frac{k_B T}{M} \le \sigma_v^2
    \end{equation}
    For a similar reason the inferred \textbf{$\sigma_{v, {\rm turb}}$ is an upper limit}. However, if you hold the turbulence fixed and assume the FWHM is an upper limit, this gives that \textbf{the temperature is a lower limit}.
}
